---
title: "Introduction to catch22"
author: "Trent Henderson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to catch22}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.height = 7,
  fig.width = 9
)
```

```{r, message = FALSE, warning = FALSE}
library(catch22)
```

`catch22` is an R package for the calculation of [22 CAnonical Time-series CHaracteristics](https://github.com/chlubba/catch22) (catch). The package is an efficient implementation that calculates time-series features coded in C.

## Core calculation function

The main function of the package is `catch22_all` which automates the computation of the 22 features. This can be run in a one-liner in R:

```{r, message = FALSE, warning = FALSE}
data <- 1 + 0.5 * 1:1000 + arima.sim(list(ma = 0.5), n = 1000)
outs <- catch22_all(data)
```

## Feature list

The full feature list can be accessed via a dataset that comes stock with the package:

```{r, message = FALSE, warning = FALSE}
View(feature_list)
```
